---
title: "Spatial Analysis"
author: "Gage Clawson"
date: "2/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rgdal)
library(sf)
library(raster)
library(tidyverse)

```

Shape files consist of four types:
  - .dbf
  - .prj 
  - .shp 
  - .shx


Read in a shapefile
```{r}
ak_shp_sf <- read_sf("shapefiles/ak_regions.shp") # specify the folder and then the shp file.


plot(ak_shp_sf[1]) #plots each column without the subset. With subset will only plot first column. 
```


Fix the problem by transforming the shape file to alaskan equal area projection

```{r}
#check crs
st_crs(ak_shp_sf)
```

```{r}
ak_shp_eq <- ak_shp_sf %>%
  st_transform(crs = 3338)  #could do the proj 4 string or an epsg code
st_crs(ak_shp_eq)

plot(ak_shp_eq[2])
```


```{r}
nrow(ak_shp_eq) #treats it as a dataframe. Wouldn't be able to do this without sf package

#geometry is sticky. Will never lose geometry from each row.


```

sf and the tidyverse
```{r}
se <- ak_shp_eq %>%
  filter(region == "Southeast")
plot(se[1])
```

## Joins

AK Population Information

```{r}
pop <- read_csv("shapefiles/alaska_population.csv")

head(pop)

```

Turn data into a simple feature

```{r}
pop_sf <- st_as_sf(pop, coords = c("lng", "lat"), crs =4326, remove = F ) #since the data is in long/lat have to read in as 4326, then we convert to alaska format
head(pop_sf)

plot(ak_shp_eq[1]) 
plot(pop_sf[1])
```

```{r}
#pop_joined_sf <- st_join(pop_sf, ak_shp_eq, join = st_within) # get an error from this because different crs

pop_sf_eq <- pop_sf %>%
  st_transform(crs = 3338)


pop_joined_sf <- st_join(pop_sf_eq, ak_shp_eq, join = st_within)

head(pop_joined_sf)
dim(pop_joined_sf)

plot(pop_joined_sf["population"]) # didn't keep the outlines of the regions 
```

Calculate total population in each region from pop_joined_sf
```{r}

#can get rid of geometry column by using st_set_geometry or as.data.frame
pop_region <- pop_joined_sf %>%
  st_set_geometry(NULL) %>%
  group_by(region_id) %>%
  summarise(total_pop = sum(population)) %>%
  ungroup()#do sum... not length

head(pop_region)
```

Add population information back to ak_shp_eq
```{r}
ak_pop_sf <- ak_shp_eq %>%
  left_join(pop_region) #probably could to full join or inner join 

head(ak_pop_sf)

plot(ak_pop_sf[4])
```

group by and summarise 

Calculate population by AK Management area
```{r}

manage_pop <- ak_pop_sf %>%
  group_by(mgmt_area) %>%
  summarise(mgmt_pop = sum(total_pop), do_union = F) #do_union = F keeps the region boundaries 

head(manage_pop)

plot(manage_pop[2])
```


Save a spatial object using write_sf
```{r}
write_sf(ak_pop_sf, "shapefiles/ak_regions_population.shp")
```













