---
title: "rasters"
author: "Gage Clawson"
date: "2/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(raster)
library(sf)
```


Read in West Coast shapefile

```{r}
wc_rgns <- read_sf("shapefiles/wc_regions_clean.shp")
head(wc_rgns)

plot(wc_rgns[1])
```

Raster data prep

Sea Surface Temp 
```{r}
r <- raster("rasters/average_annual_sst_2008.tif")
r

plot(r)
hist(r)
```

Read in a list of the SST files

```{r}
sst_files <- list.files("rasters", pattern = "average", full.names = T) #get a list of files 
sst_files
```

```{r}
sst_stack <- stack(sst_files)

sst_stack

plot(sst_stack)
```

Get the average SST

```{r}
sst_avg <- mean(sst_stack) - 273.15 #convert kelvin to celsius 
sst_avg
plot(sst_avg)
``` 

```{r}
sst_avg <- calc(sst_stack, fun = function(x){mean(x) - 273.15}, progress = "text") # progress = text helps you understand how quickly analysis is running - raster 
plot(sst_avg)

```

pipes work with raster as well

```{r}
stack(sst_files) %>%
  raster::calc(fun = function(x){mean(x) - 273.15})
``` 

Net Primary Productivity Data 

```{r}
npp <- raster("rasters/annual_npp.tif")
plot(npp)

```
```{r}
sst_avg
```

To get NPP and SST in the same projection/cellsize we are going to 
1. reproject

```{r}
#npp_proj <- projectRaster(npp, sst_avg) #reprojects and resamples

npp_proj <- projectRaster(npp, crs = ("+proj=longlat"))
plot(npp_proj)

```


2. crop

```{r}
npp_crop <- crop(npp_proj, sst_avg) #crop to match the location of sst_avg
plot(npp_crop)
```


3. resample

```{r}
npp_res <- resample(npp_crop, sst_avg, method = "ngb") #bilinear smooths the data. We want to keep as much as the initial data as possible ---> use nearest neighbor method
plot(npp_res)

##could be done in 3 lines of code ^^^
```

Analysis
```{r}
stack(npp_res, sst_avg) #see if these two layers can be stacked

```


Find cells on the WC that are good for aquaculture and those that are not

Reclassify SST

```{r}
sst_pref <- reclassify(sst_avg, c(-Inf, 12, NA, 12, 18, 1, 18, Inf, NA)) #just an easier casewhen 

plot(sst_pref)
```

```{r}

npp_pref <- reclassify(npp_res, c(-Inf, 2.6, NA, 2.6, 3.0, 1, 3.0, Inf, NA)) #just an easier casewhen 


plot(npp_pref)

```

Combine SST and NPP

```{r}
cells_pref <- npp_pref + sst_pref # really fast way to combine rasters

plot(cells_pref)


cells_pref <- overlay(sst_pref, npp_pref, fun = function(x,y){x*y}, progress = "text")

plot(cells_pref) 
plot(wc_rgns, add =T, color = NA) #run together
```

Remove international waters

```{r}
final_cells <- mask(cells_pref, wc_rgns)
plot(final_cells)

```





